An essentially algebraic formulation of dependent type theory

### 2015-02-10 Rules/rules_extension.tex
removed comma
large reorganisation
removed a comment

Using identity terms and substitution, it is also possible to define an 
extension operation on terms, which defines a term 
$\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for every $\jterm{\Gamma}{A}{x}$ and 
$\jterm{\Gamma}{\subst{x}{P}}{u}$. Thus, there is a formal definition of the
\emph{pair} $\tmext{x}{u}$. We shall not go into this, in the current paper,
but it will be very useful to have when one further develops the theory of
E-systems.

\subsubsection{Associativity of extension}
\label{comp-ee}
The inference rules asserting that extension is compatible with itself assert
that contexts are unstructured lists of type declarations. This rule is
unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{A}}{P}
    }
  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
    }
  \label{comp-ee-c}
\end{align}
Note that we may take the point of view that extension distributes over
extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
simply $Q$ since extension acts trivially at the level of families (and terms).

>>>>

Using identity terms and substitution, it is also possible to define an 
extension operation on terms, which defines a term 
$\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for every $\jterm{\Gamma}{A}{x}$ and 
$\jterm{\Gamma}{\subst{x}{P}}{u}$. Thus, there is a formal definition of the
\emph{pair} $\tmext{x}{u}$. We shall not go into this in the current paper,
but it will be very useful to have when one further develops the theory of
E-systems.

\subsubsection{Associativity of extension}
\label{comp-ee}
The inference rules asserting that extension is compatible with itself assert
that contexts are unstructured lists of type declarations. This rule is
unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{A}}{P}
    }
  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
    }
  \label{comp-ee-c}
\end{align}
